---
title: "Downloading and preparing temperature for ISIMIP 3a Baltic mizer"
author: "Max Lindmark, following Phoebe.Woodworth-Jefcoats@noaa.gov"
date: today
date-format: iso
toc: true
format: 
  html:
    page-layout: full
    embed-resources: true
knitr: 
  opts_chunk:
    fig.align: center
    out-width: 100%
editor: source
---

This document describes the workflow for preparing the ISIMIP 3a temperature data for use in the Baltic mizer model. It includes downloading and prepping the temperature data.

Temperature in mizer is to be calculated as follows:

Create a bias-corrected temperature time series as follows:

(Obsclim time series - Control data average value) + WOA average value

## Data Access

The first step was accessing the ISIMIP data. I could use the access_levante.text, but I couldn't figure out how to do it manually. I needed to do it because then I can filter by extent (Lat: 53--61, Lon: 12--32). This is necessary otherwise I get memory overload. Possibly I can use the download from the terminal if I use the `hyperfilter` function in the tidync package as I do in the plankton script --- I haven't tested.

I also used World Ocean Atlas (WOA) data as a baseline temperature. I accessed the data from the [link](https://www.ncei.noaa.gov/thredds-ocean/catalog/ncei/woa/temperature/decav/1.00/catalog.html?dataset=ncei/woa/temperature/decav/1.00/woa18_decav_t00_01.nc) in draft protocoll (using the second download option, HTTPServer: /thredds-ocean/fileServer/ncei/woa/temperature/decav/1.00/woa18_decav_t00_01.nc).

## Data downloading and wrangling

```{r}
library(tidyverse); theme_set(theme_light())
library(tidylog)
library(ncdf4)
library(lubridate)
library(raster)

home <- here::here()
```

According to the draft protocol, the temperature to put into mizer is this: (Obsclim time series - Control data average value ) + WOA average value.

First do the WOA average. Here's what I downloaded from the link in the draft Protocol

<!--![](WOA_download_1){fig-align="left"} ![](WOA_download_2){fig-align="left"} ![](WOA_download_3){fig-align="left"}-->

### Calculate the WOA average

```{r}
# https://towardsdatascience.com/how-to-crack-open-netcdf-files-in-r-and-extract-data-as-time-series-24107b70dcd

nc_dat <- nc_open(paste0(home, "/baltic/FishMIP/Temperature/DataFiles/woa18_decav_t00_01.nc"))

print(nc_dat)

attributes(nc_dat$var)
attributes(nc_dat$dim)

lat <- ncvar_get(nc_dat, "lat")
lon <- ncvar_get(nc_dat, "lon")
depth <- ncvar_get(nc_dat, "depth")

# Get the time variable. Remember: our metadata said our time units are in months since 1955-01-01 00:00:00, so you will not see a recognizable date and time format. We will take care of this later
time <- ncvar_get(nc_dat, "time")
time

# It's only length one? So this is the relevant time average, I just need to do filter and average spatially?
tunits <- ncatt_get(nc_dat, "time", "units") # Check units
tunits

months_since_1955 <- time
date_ref = as.Date("1955-01-01")
date_ref %m+% months(months_since_1955)

# Get the variable in "matrix slices"
temp_array <- ncvar_get(nc_dat, "t_an")

dim(temp_array)

temp_slice <- temp_array[ , , 1] 

image(lon, lat, temp_slice)

lonlattime <- as.matrix(expand.grid(lon, lat, depth)) # This might take several seconds

# Reshape whole temp_array
temp_long <- as.vector(temp_array)
length(temp_long)

# Create data.frame
temp <- data.frame(cbind(lonlattime, temp_long))

colnames(temp) <- c("lon", "lat", "depth", "temperature")

head(temp)

# hist(temp$depth)
# hist(temp$temperature)

# Trim the dataframe a little bit
temp <- temp %>% filter(lon > 5 & lon < 30 & lat > 50 & lat < 60)

# Now add ICES statistical subdivision information and filter
# Add ICES areas
# https://stackoverflow.com/questions/34272309/extract-shapefile-value-to-point-with-r
# https://gis.ices.dk/sf/
shape <- shapefile(paste0(home, "/baltic/FishMIP/ICES-StatRec-mapto-ICES-Areas/StatRec_map_Areas_Full_20170124.shp"))
head(shape)

pts <- SpatialPoints(cbind(temp$lon, temp$lat), 
                     proj4string = CRS(proj4string(shape)))

temp$subdiv <- over(pts, shape)$Area_27

# Rename subdivisions to the more common names and do some more filtering (by sub div and area)
sort(unique(temp$subdiv))

temp <- temp |> 
  mutate(sub_div = factor(subdiv),
         sub_div = fct_recode(subdiv,
                              "25" = "3.d.25",
                              "26" = "3.d.26",
                              "27" = "3.d.27",
                              "28" = "3.d.28.1",
                              "28" = "3.d.28.2",
                              "29" = "3.d.29",
                              "32" = "3.d.32"),
         sub_div = as.character(sub_div)) %>% 
  filter(sub_div %in% c(25, 26, 27, 28, 29, 32))

ggplot(temp, aes(lon, lat, fill = sub_div)) + 
  geom_raster() + 
  coord_sf()

# Now filter only bottom temp
sbt <- temp %>%
  drop_na(temperature) %>% 
  mutate(loc = paste(lon, lat)) %>% 
  group_by(loc) %>% 
  filter(depth == max(depth)) %>% 
  ungroup()

sbt

ggplot(sbt, aes(lon, lat, fill = temperature)) + 
  geom_raster() + 
  coord_sf()

# Calculate the mean
woa_avg_temp <- sbt %>% summarise(mean_sbt = mean(temperature))
```

Control clim, from this link: <https://data.isimip.org/search/tree/ISIMIP3a/InputData/climate/ocean/gfdl-mom6-cobalt2/ctrlclim/thetao/>

```{r}
# Read the nc file
# https://stackoverflow.com/questions/76624626/extracting-data-from-netcdf-file-in-r

nc_ctrl <- nc_open(paste0(home,
                         "/baltic/FishMIP/Temperature/DataFiles/ctrl_clim/gfdl-mom6-cobalt2_ctrlclim_thetao_15arcmin_lat53.0to61.0lon12.0to32.0_monthly_1961_2010.nc"))
print(nc_ctrl)
attributes(nc_ctrl$var)
attributes(nc_ctrl$dim)

lat <- ncvar_get(nc_ctrl, "lat")
lon <- ncvar_get(nc_ctrl, "lon")
depth <- ncvar_get(nc_ctrl, "lev")

# Get the time variable. 
time <- ncvar_get(nc_ctrl, "time")
head(time)

# Check time unit
ncatt_get(nc_ctrl, "time", "units")

# Extract time in date format
months_since_1901 <- time
date_ref = as.Date("1901-01-01")
dates <- date_ref %m+% months(months_since_1901) %>% as.character()
head(dates)

# Get the variable in "matrix slices"
ctrl_array <- ncvar_get(nc_ctrl, "thetao")
dim(ctrl_array)

# Expand grid lon lat and depth, to which be cbind the actual values
ctrl_grid <- as.matrix(expand.grid(lon, lat, depth, dates)) # This might take several seconds

# Now add in the temperatures
ctrl_df <- as.vector(ctrl_array)

ctrl_df <- data.frame(cbind(ctrl_grid, ctrl_df))

colnames(ctrl_df) <- c("lon", "lat", "depth", "date", "ctrl_temp")

# Filter NA temperatures and select the deepest depth 
ctrl_df <- ctrl_df %>% 
  drop_na(ctrl_temp) %>% 
  group_by(lon, lat) %>% 
  filter(depth == max(depth)) %>% # TODO: I wonder if this is the correct way to get the bottom depth
  ungroup() %>% 
  mutate(lon = as.numeric(lon),
         lat = as.numeric(lat),
         depth = as.numeric(depth),
         date = as.Date(date),
         ctrl_temp = as.numeric(ctrl_temp)) %>% 
  mutate(year = year(date))

# Add in ICES areas so that we can filter the domain of the mizer model
shape <- shapefile(paste0(home, "/baltic/FishMIP/ICES-StatRec-mapto-ICES-Areas/StatRec_map_Areas_Full_20170124.shp"))

pts <- SpatialPoints(cbind(ctrl_df$lon, ctrl_df$lat), 
                     proj4string = CRS(proj4string(shape)))

ctrl_df$subdiv <- over(pts, shape)$Area_27

# Rename subdivisions to the more common names and do some more filtering (by sub div and area)
sort(unique(temp$subdiv))

ctrl_df <- ctrl_df |> 
  mutate(sub_div = factor(subdiv),
         sub_div = fct_recode(subdiv,
                              "25" = "3.d.25",
                              "26" = "3.d.26",
                              "27" = "3.d.27",
                              "28" = "3.d.28.1",
                              "28" = "3.d.28.2",
                              "29" = "3.d.29",
                              "32" = "3.d.32"),
         sub_div = as.character(sub_div)) %>% 
  filter(sub_div %in% c(25, 26, 27, 28, 29, 32))

# Plot sub division
ctrl_df %>% 
  filter(date == "1961-01-01") %>% # Filter specific time
  ggplot(aes(lon, lat, fill = sub_div)) + 
  geom_raster() + 
  coord_sf()

# Test plotting temperature for the first time step
ctrl_df %>% 
  filter(date == "1961-01-01") %>% # Filter specific time
  ggplot(aes(lon, lat, fill = ctrl_temp)) + 
  geom_raster() + 
  coord_sf()

# Depth
ctrl_df %>% 
  filter(date == "1961-01-01") %>% # Filter specific time
  ggplot(aes(lon, lat, fill = depth)) + 
  geom_raster() + 
  coord_sf()

# TODO: Average by date (year + month) and plot... or do I change my dt to 1/12 and take the monthly averages?
ctrl_df %>% 
  summarise(mean_ctrl_temp = mean(ctrl_temp), .by = date) %>% 
  ggplot(aes(date, mean_ctrl_temp)) + 
  geom_line()

ctrl_annual <- ctrl_df %>% 
  summarise(mean_ctrl_temp = mean(ctrl_temp), .by = date)
```

Obs clim, from this link: <https://data.isimip.org/search/tree/ISIMIP3a/InputData/climate/ocean/gfdl-mom6-cobalt2/ctrlclim/thetao/>

```{r}
nc_obs <- nc_open(paste0(home,
                         "/baltic/FishMIP/Temperature/DataFiles/obs_clim/gfdl-mom6-cobalt2_obsclim_thetao_15arcmin_lat53.0to61.0lon12.0to32.0_monthly_1961_2010.nc"))
print(nc_obs)
attributes(nc_obs$var)
attributes(nc_obs$dim)

lat <- ncvar_get(nc_obs, "lat")
lon <- ncvar_get(nc_obs, "lon")
depth <- ncvar_get(nc_obs, "lev")

# Get the time variable. 
time <- ncvar_get(nc_obs, "time")
head(time)

# Check time unit
ncatt_get(nc_obs, "time", "units")

# Extract time in date format
months_since_1901 <- time
date_ref = as.Date("1901-01-01")
dates <- date_ref %m+% months(months_since_1901) %>% as.character()
head(dates)

# Get the variable in "matrix slices"
obs_array <- ncvar_get(nc_obs, "thetao")
dim(obs_array)

# Expand grid lon lat and depth, to which be cbind the actual values
obs_grid <- as.matrix(expand.grid(lon, lat, depth, dates)) # This might take several seconds

# Now add in the temperatures
obs_df <- as.vector(obs_array)

obs_df <- data.frame(cbind(obs_grid, obs_df))

colnames(obs_df) <- c("lon", "lat", "depth", "date", "obs_temp")

# Filter NA temperatures and select the deepest depth 
obs_df <- obs_df %>% 
  drop_na(obs_temp) %>% 
  group_by(lon, lat) %>% 
  filter(depth == max(depth)) %>% # TODO: I wonder if this is the correct way to get the bottom depth
  ungroup() %>% 
  mutate(lon = as.numeric(lon),
         lat = as.numeric(lat),
         depth = as.numeric(depth),
         date = as.Date(date),
         obs_temp = as.numeric(obs_temp)) %>% 
  mutate(year = year(date))

# Add in ICES areas so that we can filter the domain of the mizer model
shape <- shapefile(paste0(home, "/baltic/FishMIP/ICES-StatRec-mapto-ICES-Areas/StatRec_map_Areas_Full_20170124.shp"))

pts <- SpatialPoints(cbind(obs_df$lon, obs_df$lat), 
                     proj4string = CRS(proj4string(shape)))

obs_df$subdiv <- over(pts, shape)$Area_27

# Rename subdivisions to the more common names and do some more filtering (by sub div and area)
obs_df <- obs_df |> 
  mutate(sub_div = factor(subdiv),
         sub_div = fct_recode(subdiv,
                              "25" = "3.d.25",
                              "26" = "3.d.26",
                              "27" = "3.d.27",
                              "28" = "3.d.28.1",
                              "28" = "3.d.28.2",
                              "29" = "3.d.29",
                              "32" = "3.d.32"),
         sub_div = as.character(sub_div)) %>% 
  filter(sub_div %in% c(25, 26, 27, 28, 29, 32))

# Plot sub division
obs_df %>% 
  filter(date == "1961-01-01") %>% # Filter specific time
  ggplot(aes(lon, lat, fill = sub_div)) + 
  geom_raster() + 
  coord_sf()

# Test plotting temperature for the first time step
obs_df %>% 
  filter(date == "1961-01-01") %>% # Filter specific time
  ggplot(aes(lon, lat, fill = obs_temp)) + 
  geom_raster() + 
  coord_sf()

# Depth
obs_df %>% 
  filter(date == "1961-01-01") %>% # Filter specific time
  ggplot(aes(lon, lat, fill = depth)) + 
  geom_raster() + 
  coord_sf()

# TODO: Average by year and plot... or do I change my dt to 1/12 and take the monthly averages?
obs_df %>% 
  summarise(mean_obs_temp = mean(obs_temp), .by = date) %>% 
  ggplot(aes(date, mean_obs_temp)) + 
  geom_line()

obs_annual <- obs_df %>% 
  summarise(mean_obs_temp = mean(obs_temp), .by = date)
```

Compare temperature series

```{r}
temps <- bind_rows(ctrl_annual %>% rename(temp = mean_ctrl_temp) %>% mutate(type = "ctrl"),
                   obs_annual %>% rename(temp = mean_obs_temp) %>% mutate(type = "obs"))

ggplot(temps, aes(date, temp, color = type)) + 
  geom_line()
```

Phoebe’s temperature bias correction involved*:

- Create a time series of average temperature (Jan 1961 - Dec 2010) across depth ranges and over the whole model area for the different ‘realms’ (different species' vertical behavior) from the experiment data i.e. obsclim, 0.25 degree

- Create a time series of average temperature (Jan 1961 - Dec 1980) across depth ranges and over the whole model area for the different ‘realms’ from the control data i.e. ctrlclim, 0.25 degree. Calculate an average for each realm from the control time series. {note change on specifiers name, ctrclim is now named obsclim_1955-riverine-input}

- Using World Ocean Atlas climatology here, calculate the average temperature over various depth ranges and the whole model area.
Create a bias-corrected temperature time series as follows:

(Obsclim time series - Control data average value ) + WOA average value

```{r}
# Temperature time series
temp_df <- tibble(bias_corr_temp = (obs_annual$mean_obs_temp - mean(ctrl_annual$mean_ctrl_temp)) + woa_avg_temp$mean_sbt,
                  date = obs_annual$date)

ggplot(temp_df, aes(date, bias_corr_temp)) + 
  geom_line()

# Add year-averaged temperature
temp_df <- temp_df %>% 
  rename(temp = bias_corr_temp) %>% 
  mutate(year = year(date)) %>% 
  group_by(year) %>% 
  mutate(temp_agg = mean(temp)) %>% 
  ungroup()

ggplot(temp_df, aes(date, temp_agg)) + 
  geom_line()

temp_df %>% 
  distinct(year, .keep_all = TRUE) %>% 
  ggplot(aes(year, temp_agg)) + 
  geom_line()

write_csv(temp_df, paste0(home, "/baltic/FishMIP/Input-mizer/temp_df.csv"))
```
