---
title: "Downloading and preparing plankton for ISIMIP 3a Baltic mizer"
author: "Max Lindmark, following Phoebe.Woodworth-Jefcoats@noaa.gov"
date: today
date-format: iso
toc: true
format: 
  html:
    page-layout: full
    embed-resources: true
knitr: 
  opts_chunk:
    fig.align: center
    out-width: 100%
editor: source
---

This document describes the workflow for preparing the ISIMIP 3a plankton data for use in the Baltic mizer model. It includes downloading and prepping the plankton data. There are 5 groups of plankton

## Data Access
The first step was accessing the ISIMIP data. I use the code for plankton data in access_levante.txt and run it in the terminal

## Data downloading and wrangling

```{r}
library(tidyverse); theme_set(theme_light())
library(tidylog)
library(tidync)
library(ncdf4)
library(lubridate)
library(raster)
library(broom)

home <- here::here()

# Read ICES shapefile
shape <- shapefile(paste0(home, "/baltic/FishMIP/ICES-StatRec-mapto-ICES-Areas/StatRec_map_Areas_Full_20170124.shp"))
```

```{r}
# Read the nc file
# https://stackoverflow.com/questions/76624626/extracting-data-from-netcdf-file-in-r
# library(terra)
# library(tidyterra)
# https://gist.github.com/potterzot/5475ca68d2b5428044a605ce70969ef9

box <- list(lat = c(53, 61), lon = c(12, 32))

## Diatoms
t <- nc_open(paste0(home,
                       "/baltic/FishMIP/Plankton/DataFiles/gfdl-mom6-cobalt2_obsclim_phydiat-vint_15arcmin_global_monthly_1961_2010.nc"))
print(t)

diatom <- tidync(paste0(home, "/baltic/FishMIP/Plankton/DataFiles/gfdl-mom6-cobalt2_obsclim_phydiat-vint_15arcmin_global_monthly_1961_2010.nc")) %>%
  hyper_filter(
    lat = between(lat, box$lat[[1]], box$lat[[2]]),
    lon = between(lon, box$lon[[1]], box$lon[[2]])
    ) %>% hyper_tibble() %>% 
  mutate(date = as.Date("1901-01-01") %m+% months(time)) %>%
  dplyr::select(-c(time))

# Add in ICES areas so that we can filter the domain of the mizer model
pts <- SpatialPoints(cbind(diatom$lon, diatom$lat), 
                     proj4string = CRS(proj4string(shape)))

diatom$sub_div <- over(pts, shape)$Area_27

diatom <- diatom %>% 
  filter(sub_div %in% c("3.d.25", "3.d.26", "3.d.27", "3.d.28.1", "3.d.28.2", "3.d.29", "3.d.32")) %>% 
  mutate(sub_div = as.factor(sub_div)) %>% 
  droplevels() %>% 
  mutate(sub_div = fct_recode(sub_div,
                              "25" = "3.d.25",
                              "26" = "3.d.26",
                              "27" = "3.d.27",
                              "28" = "3.d.28.1",
                              "28" = "3.d.28.2",
                              "29" = "3.d.29",
                              "32" = "3.d.32"))
  


## Diazotrophs
t <- nc_open(paste0(home,
                       "/baltic/FishMIP/Plankton/DataFiles/gfdl-mom6-cobalt2_obsclim_phydiaz-vint_15arcmin_global_monthly_1961_2010.nc"))
print(t)

diaz <- tidync(paste0(home, "/baltic/FishMIP/Plankton/DataFiles/gfdl-mom6-cobalt2_obsclim_phydiaz-vint_15arcmin_global_monthly_1961_2010.nc")) %>%
  hyper_filter(
    lat = between(lat, box$lat[[1]], box$lat[[2]]),
    lon = between(lon, box$lon[[1]], box$lon[[2]])
    ) %>% hyper_tibble() %>% 
  mutate(date = as.Date("1901-01-01") %m+% months(time)) %>%
  dplyr::select(-c(time))

# Add in ICES areas so that we can filter the domain of the mizer model
pts <- SpatialPoints(cbind(diaz$lon, diaz$lat), 
                     proj4string = CRS(proj4string(shape)))

diaz$sub_div <- over(pts, shape)$Area_27

diaz <- diaz %>% 
  filter(sub_div %in% c("3.d.25", "3.d.26", "3.d.27", "3.d.28.1", "3.d.28.2", "3.d.29", "3.d.32")) %>% 
  mutate(sub_div = as.factor(sub_div)) %>% 
  droplevels() %>% 
  mutate(sub_div = fct_recode(sub_div,
                              "25" = "3.d.25",
                              "26" = "3.d.26",
                              "27" = "3.d.27",
                              "28" = "3.d.28.1",
                              "28" = "3.d.28.2",
                              "29" = "3.d.29",
                              "32" = "3.d.32"))



## Picophytoplankton
t <- nc_open(paste0(home,
                       "/baltic/FishMIP/Plankton/DataFiles/gfdl-mom6-cobalt2_obsclim_phypico-vint_15arcmin_global_monthly_1961_2010.nc"))
print(t)

pico <- tidync(paste0(home, "/baltic/FishMIP/Plankton/DataFiles/gfdl-mom6-cobalt2_obsclim_phypico-vint_15arcmin_global_monthly_1961_2010.nc")) %>%
  hyper_filter(
    lat = between(lat, box$lat[[1]], box$lat[[2]]),
    lon = between(lon, box$lon[[1]], box$lon[[2]])
    ) %>% hyper_tibble() %>% 
  mutate(date = as.Date("1901-01-01") %m+% months(time)) %>%
  dplyr::select(-c(time))

# Add in ICES areas so that we can filter the domain of the mizer model
pts <- SpatialPoints(cbind(pico$lon, pico$lat), 
                     proj4string = CRS(proj4string(shape)))

pico$sub_div <- over(pts, shape)$Area_27

pico <- pico %>% 
  filter(sub_div %in% c("3.d.25", "3.d.26", "3.d.27", "3.d.28.1", "3.d.28.2", "3.d.29", "3.d.32")) %>% 
  mutate(sub_div = as.factor(sub_div)) %>% 
  droplevels() %>% 
  mutate(sub_div = fct_recode(sub_div,
                              "25" = "3.d.25",
                              "26" = "3.d.26",
                              "27" = "3.d.27",
                              "28" = "3.d.28.1",
                              "28" = "3.d.28.2",
                              "29" = "3.d.29",
                              "32" = "3.d.32"))


## Mesozooplankton
t <- nc_open(paste0(home,
                       "/baltic/FishMIP/Plankton/DataFiles/gfdl-mom6-cobalt2_obsclim_zmeso-vint_15arcmin_global_monthly_1961_2010.nc"))
print(t)

meso <- tidync(paste0(home, "/baltic/FishMIP/Plankton/DataFiles/gfdl-mom6-cobalt2_obsclim_zmeso-vint_15arcmin_global_monthly_1961_2010.nc")) %>%
  hyper_filter(
    lat = between(lat, box$lat[[1]], box$lat[[2]]),
    lon = between(lon, box$lon[[1]], box$lon[[2]])
    ) %>% hyper_tibble() %>% 
  mutate(date = as.Date("1901-01-01") %m+% months(time)) %>%
  dplyr::select(-c(time))

# Add in ICES areas so that we can filter the domain of the mizer model
pts <- SpatialPoints(cbind(meso$lon, meso$lat), 
                     proj4string = CRS(proj4string(shape)))

meso$sub_div <- over(pts, shape)$Area_27

meso <- meso %>% 
  filter(sub_div %in% c("3.d.25", "3.d.26", "3.d.27", "3.d.28.1", "3.d.28.2", "3.d.29", "3.d.32")) %>% 
  mutate(sub_div = as.factor(sub_div)) %>% 
  droplevels() %>% 
  mutate(sub_div = fct_recode(sub_div,
                              "25" = "3.d.25",
                              "26" = "3.d.26",
                              "27" = "3.d.27",
                              "28" = "3.d.28.1",
                              "28" = "3.d.28.2",
                              "29" = "3.d.29",
                              "32" = "3.d.32"))


## Microzooplankton
t <- nc_open(paste0(home,
                       "/baltic/FishMIP/Plankton/DataFiles/gfdl-mom6-cobalt2_obsclim_zmicro-vint_15arcmin_global_monthly_1961_2010.nc"))
print(t)

micro <- tidync(paste0(home, "/baltic/FishMIP/Plankton/DataFiles/gfdl-mom6-cobalt2_obsclim_zmicro-vint_15arcmin_global_monthly_1961_2010.nc")) %>%
  hyper_filter(
    lat = between(lat, box$lat[[1]], box$lat[[2]]),
    lon = between(lon, box$lon[[1]], box$lon[[2]])
    ) %>% hyper_tibble() %>% 
  mutate(date = as.Date("1901-01-01") %m+% months(time)) %>%
  dplyr::select(-c(time))

# Add in ICES areas so that we can filter the domain of the mizer model
pts <- SpatialPoints(cbind(micro$lon, micro$lat), 
                     proj4string = CRS(proj4string(shape)))

micro$sub_div <- over(pts, shape)$Area_27

micro <- micro %>% 
  filter(sub_div %in% c("3.d.25", "3.d.26", "3.d.27", "3.d.28.1", "3.d.28.2", "3.d.29", "3.d.32")) %>% 
  mutate(sub_div = as.factor(sub_div)) %>% 
  droplevels() %>% 
  mutate(sub_div = fct_recode(sub_div,
                              "25" = "3.d.25",
                              "26" = "3.d.26",
                              "27" = "3.d.27",
                              "28" = "3.d.28.1",
                              "28" = "3.d.28.2",
                              "29" = "3.d.29",
                              "32" = "3.d.32"))

# Combine all into a dataframe
pl <- bind_rows(diatom %>% rename(mole_content = `phydiat-vint`) %>% mutate(type = "diatom"),
                diaz %>% rename(mole_content = `phydiaz-vint`) %>% mutate(type = "diazotrophs"),
                pico %>% rename(mole_content = `phypico-vint`) %>% mutate(type = "phypico"),
                meso %>% rename(mole_content = `zmeso-vint`) %>% mutate(type = "mesozp"),
                micro %>% rename(mole_content = `zmicro-vint`) %>% mutate(type = "microzp"))
  
# Do the same conversions that Phoebe does in the script Plankton_ts.jnl
# ! We want total carbon in g, so we'll need to:
# ! Convert from mol C to g C (*12.011)
# ! Convert area from km^2 to m^2 (*10^6)
# ! Sum 

pl$g_C <- pl$mole_content*12.011

# TODO: check if this is correct. The nc files are in units: mol m-2, so we convert from k^m to m^2??
# I think I need to convert to my area, right? 2.49×10^11 m^2
# But how can I do that without knowing the area of each "grid cell" the point represents?
# Do I sum across spatial points and then just... multiply by areal extent?
# But if it's in unit per m2, then I ought to take an average...

pl_sum <- pl %>% 
  summarise(g_C = sum(g_C), .by = c(date, type)) %>% 
  mutate(sum_g_C = g_C*2.49e+11)
```

Convert the units to get information about the intercept and slope of the resource size-spectra. This is following Phoebe's code Prep_Plankton_therMizer.Rmd, which reads in data prepared in Plankton_ts.jnl, which reads in the nc files and "plankton carbon densities were converted to total carbon and summed over the model domain". Now we need to make further conversions.

Here, this total carbon is used to create resource spectra at each time step.

Size class ranges, for reference (ESD = Equivalent Spherical Diameter):  
- Phyto = 0.2 - 200 um ESD (mid-point size = 100.1 um ESD)  
-	pico = 0.2 - 10 um ESD (mid-point size = 5.1 um ESD)  
-	large (diatoms and diazotrophs) = 10 - 200 um ESD (mid-point size = 105 um ESD)  
- Zoo = 2 - 20,00 um ESD (based on the literature)(mid-point size = 10001 um ESD)  
-	zmicro = 2 - 200 um ESD (mid-point size = 101 um ESD)  
-	zmeso = 200 - 20000 um ESD (mid-point size = 10100 um ESD)  

These size classes were informed by Dunne et al. [2005](https://doi.org/10.1029/2004GB002390), [2012](https://doi.org/10.1029/2010GB003935), and [2013](https://doi.org/10.1175/JCLI-D-12-00150.1), [Liu et al. 2021](https://doi.org/10.1029/2021GL094367), and [Stock et al. 2020](https://doi.org/10.1029/2019MS002043).

Conversions used:  
- Convert g C to gww --> $\times 10$  
- Convert um ESD to gww --> $\frac{4}{3}\pi(0.5\times0.0001\times size)^3$

```{r}
# Create variables for referencing the size class mid points, in gww
pico_mid <- (4/3)*pi*((0.5*0.0001*5.1)^3)
large_mid <- (4/3)*pi*((0.5*0.0001*105)^3) 
micro_mid <- (4/3)*pi*((0.5*0.0001*101)^3)
meso_mid <- (4/3)*pi*((0.5*0.0001*10100)^3)

mid_df <- data.frame(type = c("diatom",  "diazotrophs", "phypico", "mesozp", "microzp"),
                     midpoint = c(large_mid, large_mid,     pico_mid,  pico_mid, micro_mid))

# Convert total carbon to gww and then get numerical abundance by dividing by size class mid point
# This step assumes that all plankton are the midpoint size
pl <- pl %>% 
  left_join(mid_df, by = "type") %>% 
  mutate(abund = g_C*10/midpoint) # TODO: Pheobes g_C is scaled by to km^2, mine only converted to carbon


# TODO: I don't need spatial data to estimate the intercept and slope, but at this point I'm not sure how to sum, so I calculate the mean in the spatial domain.
pl_fit <- pl %>% 
  summarise(mean_abund = mean(abund), .by = c(date, type)) %>% 
  left_join(mid_df)

ggplot(pl_fit, aes(midpoint, log(mean_abund), color = as.numeric(date))) + 
  geom_jitter() +
  geom_line()

# Fit a model to estimate the slope and intercept for all time steps
m <- lm(log10(mean_abund) ~ 0 + log10(midpoint)*as.factor(date), data = pl_fit)

# Main effect of log(midpoint)
main_slope <- tidy(m) %>%
  filter(term == "log10(midpoint)") %>%
  distinct(estimate)

# Extract the slopes at each time step
slopes <- tidy(m) %>% 
  filter(grepl("log10", term)) %>% 
  filter(!term == "log10(midpoint)") %>% 
  mutate(slope = estimate + main_slope$estimate) %>% 
  separate(term, into = c("term", "date"), ":") %>% 
  mutate(date = str_replace(date, "as.factor", "")) %>% 
  mutate(date = str_replace(date, "(date)", "")) %>% 
  mutate(date = str_replace(date, "[()]", "")) %>% 
  mutate(date = str_replace(date, "[()]", "")) %>% 
  mutate(year = year(as.Date(date))) %>% 
  dplyr::select(term, year, date, slope) %>% 
  rename(estimate = slope) %>% 
  mutate(term = "slope")

ggplot(slopes, aes(estimate)) +
  geom_histogram()

ggplot(slopes, aes(as.Date(date), estimate)) +
  geom_point()

slopes %>% 
  filter(year < 1980) %>% 
  ggplot(aes(as.Date(date), estimate)) +
  geom_point() + 
  geom_line() + 
  geom_hline(yintercept = -2-0.8+(2/3), linetype = 2)

# Very cyclical! And also... the value is quite different from the size-spectrum slope I currently have, 

# Now get the intercepts
inter <- tidy(m) %>% 
  filter(!grepl(":", term)) %>% 
  filter(!term == "log10(midpoint)") %>% 
  separate(term, into = c("term", "date"), "[)]") %>% 
  mutate(term = "intercept",
         year = year(as.Date(date))) %>% 
  dplyr::select(term, year, date, estimate) %>% 
  mutate(estimate = 10^estimate)
  

coefs <- bind_rows(inter, slopes) %>% 
  mutate(month = month(as.Date(date)))

# Add the scaled estimates...
coefs <- coefs %>% 
  group_by(term) %>% 
  mutate(estimate_sc = estimate / mean(estimate)) %>% 
  ungroup()

ggplot(coefs, aes(as.Date(date), estimate_sc, color = factor(month))) +
  geom_point() + 
  facet_wrap(~term, ncol = 1)

ggplot(coefs, aes(as.Date(date), estimate_sc, color = factor(month))) +
  geom_point() + 
  facet_wrap(~term, ncol = 1, scales = "free")

# Huge difference between months... perhaps I need to aggregate?
coefs <- coefs %>% 
  group_by(term, year) %>% 
  mutate(estimate_agg = mean(estimate)) %>% 
  group_by(term) %>% 
  mutate(estimate_agg_sc = estimate_agg / mean(estimate_agg)) %>% 
  ungroup()
  
ggplot(coefs, aes(as.Date(date), estimate_agg_sc, color = factor(month))) +
  geom_point() + 
  facet_wrap(~term, ncol = 1, scales = "free")
```

Save

```{r}
# Make it wide before saving... much easier later
coefs_save <- coefs %>% 
  filter(term == "intercept") %>% 
  rename(intercept = estimate_agg_sc) %>% 
  dplyr::select(date, intercept)

coefs_save <- coefs_save %>% 
  left_join(coefs %>% 
              filter(term == "slope") %>% 
              rename(slope = estimate_agg_sc) %>% 
              dplyr::select(date, slope), by = "date")

# Erm, because there's a global intercept we don't have the slope for the first iteration
# But since it's all aggregated by year anyway I can just fill it in now
coefs_save
coefs_save[1, 3] <- coefs_save[2, 3]
coefs_save

write_csv(coefs_save, paste0(home, "/baltic/FishMIP/Input-mizer/resource_df.csv"))
```
